Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-10-14 exp1/present2013.tex
remove comments
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@673 43ef1190-90fd-4dd7-a769-d0121f02b80e
some cuts
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@674 43ef1190-90fd-4dd7-a769-d0121f02b80e

%\documentclass[english,aspectratio=196]]{ifislide}
%\documentclass[english,print,aspectratio=169]{ifislide}
\documentclass[english,handout,aspectratio=169]{ifislide}
\usepackage{verbatim}
\usepackage{wrapfig}
\setbeameroption{show notes}
\usepackage[utf8]{inputenc}
% Packages used specific for this lecture:

%\usepackage{bussproofs}
%\usepackage{textpos}

%\EnableBpAbbreviations
%% \usepackage{tikz}

%% \usetikzlibrary{arrows,shapes,mindmap}
%% \usetikzlibrary{decorations.pathmorphing}
%% \tikzstyle{resource}=[ellipse,draw=black!40,fill=blue!20,thick, font=\ttfamily]
%% \tikzstyle{blank}=[ellipse,draw=black!40,fill=white!40,thick, font=\ttfamily]
%% \tikzstyle{literal}=[rectangle,text centered,draw=black!40,fill=black!20,thick]
%% \tikzstyle{predicate}=[sloped,above,text centered,midway,font={\scriptsize\ttfamily}]
%% \tikzstyle{predicateb}=[left,midway,font={\scriptsize\ttfamily}]
%% \tikzstyle{class}=[rectangle,text centered,draw=black!40,fill=white,thick]
%% \usetikzlibrary{arrows,automata}

@@

% set lecture number, date, title:
%\LectureNumber{13}
\LectureDate{24.10.2013}

@@

%\IfiTitleSlide     % inserts title frame

%\IfiTOCSlide       % inserts today's plan frame

@@

\begin{frame}{Not all benchmarks are created equal}

  \begin{itemize}
  \item Natural benchmarks.
  \item Benchmarking with broad and well-known workloads. 
  \item Artificial use-case driven benchmarks.
  \end{itemize}

\end{frame}
\note{
  \begin{itemize}
  \item Natural benchmarks, query log-derived, DBPedia SPARQL
    Benchmark
  \item If your queries are all fully known, by all means, run them
  \item In most scientific relevant cases, no such assumptions can be
    made
  \item What I will present will likely create artifical workloads
  \item Use-case driven will seem artificial for everything else
  \end{itemize}
}

@@

\begin{itemize}
\item Large number of parameters to test for complex scenarios.
\item No structured approach to investigate flaws.
\item No meaningful summary of the overall performance.
\item Standardized benchmarks cannot test assertions outside of the
  standard.
\item Attempting to neutralize the effect of certain optimization
  techniques oversimplifies the test.
\end{itemize}
\end{frame}
\note{  
  \begin{itemize}
  \item If you standardize hardware, you can't test claims about
  improved hardware. You'd need interactions too
  \item Cache-buster is evil
\end{itemize}

@@

  \begin{itemize}
  \item Pioneered by Fischer in the 1930s.
  \item Well-established in many fields of engineering, agriculture,
    medicine.
  \item Well-suited to manage very complex experiments.
  \item Requires parameterization of experiments.
  \end{itemize}
\end{frame}
\note{How many is familiar with DoE?}

@@

  \begin{itemize}
  \item Interaction effects may be complex, like one implementation is
    better on CONSTRUCT queries, on small datasets
\item We still need to understand which effects are important and
  significant, we do that in the discussion of our experiment
\end{itemize}
}
% SECTION

@@

@@

  \begin{itemize}
  \item Implemented in R, using packages \texttt{DoE.base} and
    \texttt{FrF2}.
  \item Two experiment machines, one machine running queries
  \item No warm-up, no cache busting
  \item 256 experiments in full factorial, no replications, compare
    favorably to benchmarks
  \end{itemize}
}

@@

\note{
  \begin{itemize}
  \item Unreplicated, so we don't have a standard error, so Lenth used
    a restricted subset the median of the effects to create a Pseudo
    Standard Error
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
    to that many more triples need to be searched for a language tag
    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''?
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine''
  \end{itemize}
}

\begin{frame}{Hypothesis tests}

@@

\item May create a one-sided two-sample t-test with 4 values for each of the levels.
\item Lends support to $H_1$ with a high probability, $p=1.16 \cdot
10^{-07}$.

@@

  \begin{itemize}
  \item Find most of the same significant effects as the full experiment.
  \item However, we see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$.
  \item ``Machine'', ``TripleC'' and ``Range'' are inactive, we can
    perform hypothesis test as above, and find $p = 5.9 \cdot 10^{-6}$.
  \end{itemize}
\end{frame}
\note{
  \begin{itemize}
  \item We see that ``Implement:TripleC'' emerges only for
    $\alpha=0.15$, which is important to better in real studies.
  \item We have shown that how the experiment could be set up, we have
    shown the analysis and a hypothesis test.
  \item Should this simplistic experiment stand up to scrutiny?
  \end{itemize}
}

@@

\begin{frame}{Linearity and choice of levels}
  \begin{itemize}
  \item 2-level experiments can only model linear effects.
  \item Choice of levels may be critical.
  \item If response is known to be non-linear, using more levels must be considered.
  \end{itemize}

\end{frame}
\note{2-level good enough for identifying effects, insufficient as
  statistical model}

>>>>

%\documentclass[english,aspectratio=196]]{ifislide}
%\documentclass[english,print,aspectratio=169]{ifislide}
\documentclass[english,handout,aspectratio=169]{ifislide}
\usepackage{verbatim}

%\setbeameroption{show only notes}
\usepackage[utf8]{inputenc}

@@

\LectureDate{24.10.2013}

@@

@@

@@

\begin{itemize}
\item Large number of parameters to test for complex scenarios.
\item No structured approach to investigate flaws.
\item No meaningful summary of the overall performance.
\item Standardized benchmarks cannot test assertions outside of the
  standard.
\item Attempting to neutralize the effect of certain optimization
  techniques oversimplifies the test.
\end{itemize}
\end{frame}
\note{  
  \begin{itemize}
  \item If you standardize hardware, you can't test claims about
  improved hardware. You'd need interactions too
  \item Caching is a good practice, and its effect need to be a part
    of the test, not gotten rid of.
  \item That said benchmarking is still nice to have if you know what
    queries you will run, it can't replace that.
\end{itemize}

@@

  \begin{itemize}
  \item Pioneered by Fischer in the 1920s.
  \item Well-established in many fields of engineering, agriculture,
    medicine.
  \item Well-suited to manage very complex experiments.
  \item Requires parameterization of experiments.
  \end{itemize}
\end{frame}
\note{How many is familiar with DoE?}

@@

  \begin{itemize}
  \item Interaction effects may be complex, like one implementation is
    better on CONSTRUCT queries, on small datasets
\item We still need to understand which effects are important and
  significant, we do that in the discussion of our experiment
\end{itemize}
}

% SECTION

@@

\end{frame}
\note{Go through very quickly}

@@

\note{Implemented in R, system for statistics.}

@@

\note{
  \begin{itemize}
  \item Unreplicated, so we don't have a standard error, so Lenth used
    a restricted subset the median of the effects to create a Pseudo
    Standard Error
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, 
  \item Interactions with Implement with joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
    to that many more triples need to be searched for a language tag
    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''? Leave that hanging for
    a bit.
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine''
  \end{itemize}
}

\begin{frame}{Preparing hypothesis tests}

@@

\item May create a one-sided two-sample t-test with 4 values for each
  of the levels of ``Implement''.
\item Lends support to $H_1$ with a high probability, $p=1.16 \cdot
10^{-7}$.

@@

  \begin{itemize}
  \item Find most of the same significant effects as the full experiment.
  \item However, we see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$.
  \item ``Machine'', ``TripleC'' and ``Range'' are inactive, we can
    perform hypothesis test as above, and find $p = 5.9 \cdot 10^{-6}$.
  \end{itemize}
\end{frame}
\note{
  \begin{itemize}
  \item We see that ``Implement:TripleC'' emerges only for
    $\alpha=0.15$, which is important to improve when designing a
    study for actual tests.
  \item We have shown that how the experiment could be set up, we have
    shown the analysis and a hypothesis test.
  \item Should this simplistic experiment stand up to scrutiny?
  \end{itemize}
}

@@
